19Q1

### 2019-05-10 19Q1.Rmd
updated Rmd

```{r libraries}
library(circlize)
library(ComplexHeatmap)
library(cowplot)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(scales)
library(tidyverse)
library(VennDiagram)

@@

```{r g2p_interpretations}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p[, 1:2] <- NULL
g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p <- subset(g2p, grepl("^CID", g2p$DrugID))
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p <- merge(g2p, lineage_converter, by = "grouped_general_doid", all.x = TRUE)
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))

g2p_dgls <-  unique(g2p$Drug_Gene_Lin[!is.na(g2p$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p$Drug_Gene[!is.na(g2p$Drug_Gene)])
g2p_drugs <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[1]))
g2p_genes <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[2]))
g2p_lins <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[3]))
g2p_druggenes
g2p_drugs
g2p_genes
g2p_lins

@@

```{r testing, eval=FALSE}
unique(g2p$group_general_lineage_name)
levels(g2p$group_general_lineage_name)

>>>>

```{r libraries}
library(circlize)
library(ComplexHeatmap)
library(cowplot)
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(jsonlite)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(scales)
library(stringr)
library(tidyverse)
library(VennDiagram)

@@

```{r g2p_parsing_functions}
ExtractBiomarkerType <- function(json_string) {
  json_string <- as.character(json_string)
  biomarker_type <- ((str_extract(string = json_string, pattern = "biomarker_type.+") %>% strsplit(split = ","))[[1]][1] %>% str_extract(pattern = "u'.+") %>% str_match("'(.*?)'"))[,2]
  return(biomarker_type)
}
```

```{r g2p_parsing, eval=FALSE}
# g2p_json <- "./../htscreens_giant_files/g2p_all.json"
# g2p_all <- fromJSON(sprintf("[%s]", paste(readLines(g2p_json), collapse = ",")), flatten = TRUE)
# saveRDS(g2p_all, "./../htscreens_giant_files/g2p_all_flattened_json.rds", compress = "xz")

g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p[, 1:2] <- NULL
g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p <- subset(g2p, grepl("^CID", g2p$DrugID))
g2p$Direction <- ifelse(g2p$Direction == 1 | g2p$Direction == "", NA, as.character(g2p$Direction))
g2p$Gene <- ifelse(is.na(g2p$Gene) & g2p$Drug == "daunorubicin", "MLL", as.character(g2p$Gene))
g2p <- g2p[, c("Drug", "DrugID", "Direction", "Gene", "Ref", "Alt", "Start", "End", "Chromosome", "Feature.Names", "Mutation", "JSON", "Evidence.Level", "Phenotype.Description", "Phenotype.Family", "Phenotype.ID", "Sequence.ID", "grouped_general_doid")]
g2p <- add_column(g2p, biomarker_type = tolower(sapply(g2p$JSON, ExtractBiomarkerType, simplify = TRUE)), .after = "Feature.Names")
g2p <- add_column(g2p, biomarker_type_curated = as.character(894), .after = "biomarker_type")
g2p$biomarker_type_curated <-
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = ",")), NA,
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "del")), "deletion",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "ins|dup")), "insertion",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "oncogenic mutation|mutant")), "mutation",
  ifelse(g2p$Feature.Names == "KRAS G12fs*3", "deletion",
  ifelse(g2p$biomarker_type == "intron variant", "intron variant",
  ifelse(g2p$biomarker_type == "microsatellite", "microsatellite",
  ifelse(g2p$biomarker_type == "positive", "positive",
  ifelse(g2p$biomarker_type == "rearrange", "rearrangement",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "fusion|FUSION|Fusion")), "fusion", 
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "[A-Z0-9]-[A-Z0-9]")), "fusion",
  ifelse(g2p$biomarker_type == "synonymous", "synonymous",
  ifelse(g2p$biomarker_type == "biallelic inactivation", "biallelic inactivation",       
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "[A-Z][0-9]+[A-WYZ]")), "missense",
  ifelse(!is.na(str_match(string = g2p$Feature.Names, pattern = "[A-Z][0-9]+([X]|[*])")), "nonsense", NA)))))))))))))))
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p <- merge(lineage_converter, g2p, by = "grouped_general_doid", all.y = TRUE)
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))

write.csv(g2p, "./data_munging/g2p_db_edited.csv")

g2p_dgls <-  unique (g2p$Drug_Gene_Lin[!is.na(g2p$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p$Drug_Gene[!is.na(g2p$Drug_Gene)])
g2p_drugs <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[1]))
g2p_genes <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[2]))
g2p_lins <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[3]))
g2p_druggenes
g2p_drugs
g2p_genes
g2p_lins

!is.na(str_match(string = "ALK L1198F", pattern = "[A-Z][0-9]+[A-WYZ]"))

@@

```{r g2p}
g2p <- read.delim("./data_munging/g2p_db_edited.csv", sep = ",", header = TRUE)

g2p_dgls <-  unique (g2p$Drug_Gene_Lin[!is.na(g2p$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p$Drug_Gene[!is.na(g2p$Drug_Gene)])
g2p_drugs <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[1]))
g2p_genes <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[2]))
g2p_lins <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[3]))
```

```{r testing, eval=FALSE}
g2p$JSON[1]
test <- filter(g2p, is.na(Gene))
test$JSON[2]

unique(g2p$group_general_lineage_name)
levels(g2p$group_general_lineage_name)
